### 2024-01-20 outputs/paper/paper.qmd
Introduction added and refereces.bib updated accordingly. Introduction section has been cited properly at places where it was necessary.

---
title: "An Analysis of a Rising Trend in Dollar Loss in Fire Incidents With and Without Sprinkler Systems in Toronto over a decade (2011-2022)"
author: 
  - Aryaman Sharma
thanks: "Code and data are available at: https://github.com/Ary4m3n/fire-incidents.git"
date: today
date-format: long
abstract: "A Fire Incidents dataset was used to assess the general trend in dollar loss in fire incidents and the impact of existence of sprinkler systems installed across the city of Toronto. Based on the analysis, a general rise over the years in the dollar loss in fire incidents as well as a massive amount of loss to property was found wherever sprinkler systems were not installed. This paper suggests that an installation of sprinkler systems will help reduce the impact of fire incidents on loss to property in Toronto."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction
The Great Fire of Toronto in 1904 caused mass destruction. The fire demolished around 20 acres of the industrial area in the city of Toronto, demolishing at least 98 buildings, causing a dollar loss of $10 million and leaving numerous thousands unemployed [@greatFire]. Since then we have seen an improvement in city codes and policies, better civil planning and extensive training of firefighters [@fireMedium]. Till date, the Toronto Fire Services report approximately 10,000 fires each year which is of immense concern to the city of Toronto and its residents [@fireData].

The Fire Incidents data of Toronto includes incident reports of fire incidents as defined by the Ontario Fire Marshal from 2011-2022 [@openData]. This paper delves deeper into the trend of estimated dollar loss in fire incidents in the city of Toronto from 2011-2022 and also analyses the impact of an installation of sprinkler systems on the dollar loss. The aim of the paper is to analyse the impact of better planning, specifically an installation of sprinkler systems, on the extent of loss faced by the residents of Toronto.

In this paper, Fire Incidents data [@openData] was used to first explore if there is a trend in the estimated dollar loss over the years, and then was used to find any correlation between better civil planning, i.e. the installation of sprinkler systems and a lesser loss of property. It was found that there is a general rising trend in the dollar loss between 2011 and 2022. The paper also found that the dollar loss was drastically higher for fire incidents where there was no sprinkler system installed. This leads us to make an important judgement that better civil planning helps reduce the impact of such unfortunate events.

This paper is structured using the following sections: Data, Results and Discussion In the Data (@sec-data) section, the data source of the dataset from Open Data Toronto [@openData] is discussed and the data cleaning process is outlined. In the Results (@sec-results) section, the paper summarizes the data findings and relevant graphs of the trends observed. The paper ends with the Discussion (@sec-discussion) section, where the findings of the paper have been analysed and delved deeper into, and a further scope for the paper has been discussed.

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References